\LARGE \bf Expectation-Maximization based defense mechanism for distributed Model Predictive Control

### 2022-02-03 article/article.tex
change structure and add some comments

\loop\iftrue%
\errmessage{This article is copyrighted by ___ and should not be TeXed}%
\repeat%

@@

\section{INTRODUCTION}
% % % \mpc{} \dmpc{}

@@

\begin{equation}
  \label{eq:constraint}
  \sum^{\nsubsystems}_{i=1}\Gamma_{i}\uik\preceq\umax
\end{equation}
where ${\Gamma_{i}:\R^{n_{u}\times n_{u}}}$, ${\umax:\R^{n_{u}\times 1}}$ and ${n_{u}=\card{\uik}}$.

\begin{problem}{Global \mpc{} Problem.}\label{Pb:GOP}
\begin{equation*}
\begin{matrix}

\underset{\uiseq}{\mathrm{minimize}}&\resizebox{0.35\textwidth}{!}{$\overbrace{\sum\limits^{M}_{i=1} \overbrace{\sum_{j=1}^{\predhorz} \norm{\mpcvec{v}[i][k+j][k]}^{2}_{Q_i}+\norm{\mpcvec{u}[i][k+j-1][k]}^{2}_{R_i}}^{\textstyle{} \obji}}^{\textstyle{} \globobj}$}\\
\mathrm{subject~ to}&~\eqref{eq:systems},\eqref{eq:upositive}\ \mathrm{and}\ \eqref{eq:constraint}
\left\}\small
\begin{aligned}
  &\forall i\in \{1:\nsubsystems\}\\
  &\forall j\in \{1:\predhorz\}
\end{aligned}\right.

@@

$\optglobobj$ denotes the optimal value of the problem~\ref{Pb:GOP}, and ${\optuiseq}$ is the optimal control sequences.
the problem~\ref{Pb:GOP} is solved at each time $k$, and the ${\optuikk}$ are applied in each respective $i$ subsystem, following a receding horizon strategy.
\end{problem}

\subsection{Distributed Model Predictive Control}\label{ssec:dMPC}
\todo{\dmpc{},
}
\begin{subequations}
  \begin{equation}
    \left.
      \small
      \begin{aligned}
        J_{i}^{\star}(\thetaik)&=\underset{\uiseq}{\mathrm{minimize}} \obji\\
        \mathrm{s.t.} &\quad\eqref{eq:systems}\ \&\ \eqref{eq:upositive}\\
        &\quad\Gamma_{i}\uik\preceq\thetaik:\lambdaik\\
      \end{aligned}
    \right\}
    \small
    \begin{aligned}
      &\forall i\in \{1:\nsubsystems\}\\
      &\forall j\in \{1:\predhorz\}
    \end{aligned}
    \label{eq:DOP_local}
  \end{equation}
  \begin{equation}
    \small
    \begin{aligned}
      J^{\star}&=\underset{\thetaiseq}{\mathrm{minimize}}\sum^{\nsubsystems}_{i=1} J^{\star}_i(\thetaik)\\
      \mathrm{s.t.} &\quad \sum_{i=1}^{\nsubsystems}\thetaik\preceq\umax
    \end{aligned}
    \label{eq:DOP_master}
  \end{equation}
\end{subequations}

@@

\todo{The sum $\sum_{i=1}^{M}\vec{\theta}_{i}$ can also be represented by the matrix multiplication $I_{c}^{M}\vec{\theta}$. Where ${c=\pi_{\vec{u}_{i}(k:k+N_{p}-1|k)}=N_{p}\pi_{\uik}}$.}

\begin{equation}
  \label{eq:thetaNegot}
\thetai\pplusone=\thetai\p+\rho\left(\vec{\lambda}_{i}^{\star} (\vec{\theta}_{i}\p)-I_{c}^{M} \pseudoinv{I_{c}^{M}}\vec{\lambda}^{\star} (\vec{\theta}\p)\right)
\end{equation}
\todo[verify if need to use max]{
\begin{equation}
  % \label{eq:thetaNegot}
  \thetai\pplusone=
  \left\{ \thetai\p+\rho\left(\vec{\lambda}_{i}^{\star} (\vec{\theta}_{i}\p)-I_{c}^{M} \pseudoinv{I_{c}^{M}}\vec{\lambda}^{\star} (\vec{\theta}\p)\right)
  ,0\right\}
\end{equation}
}
\no{\todo[delete probably]{
  But observe in~\eqref{eq:projectedSubgradient} that if ${\vec{\theta}_{i}\p=\0}$, ${(\vec{\lambda}_{i}^{\star} (\vec{\theta}_{i}\p)-I_{c}^{M} \pseudoinv{I_{c}^{M}}\vec{\lambda}^{\star} (\vec{\theta}\p))}$ needs to be negative so $\vec{\theta}_{i}\pplusone$ becomes negative, which is impossible \question{is it though?} since $\lambdaik\succeq\0$, from the KKT optimality conditions, seen in~\cite[Chapter 5]{BoydVandenberghe2004}.
}}

@@

\todo{As seen in~\cite{BemporadEtAl2002,BorrelliEtAl2017}, the solution for a constrained linear quadratic regulation problem yields in a \pwa{} function.
Similarly, we can get an explicit  solution for its dual variables $\vec{\lambda}_{i}$, which is also a \pwa{} function with respect to $\vec{\theta}_{i}$:}
\begin{equation}
  \begin{aligned}
    \label{eq:lambdafuntheta}
    \lambdai=
    \begin{cases}
      -P_{i}^{1}\vec{\theta}_{i}-\vec{s}_{i}^{1}[k]&\text{if}\ G_{1}[k]\thetai \preceq z_{1}[k] \\
      \qquad\quad \vdots&\qquad\quad \vdots\\
      -P_{i}^{N}\vec{\theta}_{i}-\vec{s}_{i}^{N}[k]&\text{if}\ G_{N}[k]\thetai \preceq z_{N}[k] \\
    \end{cases}
  \end{aligned}
\end{equation}
The halfspaces defined by the set of pairs $(G_{j}[k],z_{j}[k])$ represent a combination of active constraints for a given time $k$.

\begin{challenge}
It is important to note that since the actual state is unknown we cannot anticipate how the space is partitioned, nor the values of $\vec{s}_{i}[k]$.
\end{challenge}
 As each constraint can be active or inactive, for a group of $n$ constraints, \todo{potentially} we can have $N=2^{n}$ different combinations of active/inactive sets, consequently $N$ regions (or zones).
\todo{
  Here we say potentially because depending on the constraints, some of these combinations can be infeasible.\question{Remove??}
}

@@

\subsection{Detection and mitigation}
\begin{assumption}
  $\gamma_{i}(\cdot)$ is the same during the negotiation phase for a given time $k$ (it does not depend on $p$).
\end{assumption}

\begin{assumption}
  We suppose the agent chooses a linear function such as $\tilde{\lambdai}=\gamma_{i}(\vec{x})=\Tik\vec{x}$ resulting in:
  \begin{equation}
    \label{eq:linear_cheating}
    \tilde{\lambdai}=
        \begin{cases}
      -\Tik P_{i}^{1}\vec{\theta}_{i}-\Tik \vec{s}_{i}^{1}[k]&\text{if}\ G_{1}\thetai \preceq z_{1} \\
      \qquad\quad \vdots&\qquad\quad \vdots\\
      -\Tik P_{i}^{N}\vec{\theta}_{i}-\Tik \vec{s}_{i}^{N}[k]&\text{if}\ G_{N}\thetai \preceq z_{N} \\
    \end{cases}
  \end{equation}
and we define $\tilde{P}_{i}^{j}[k]=\Tik P_{i}^{j}$ and $\tilde{\vec{s}}_{i}^{j}[k]=\Tik\vec{s}_{i}^{j}[k]$.
\end{assumption}
\todo{
  Since $\Theta_{i}$ are diagonal we can assure\pdfnote{proof?} there exists one \question{zone?}zone where all constraints are active. We will fix $j=1$ for this zone and called it the $1$-zones.
  We can estimate the values for this partition using:
}
\begin{equation}
  \label{eq:lambdafuntheta_tilde}
\tilde{\lambdai}=\gamma_{i}(\lambdai(\thetai))=-\widehat{\tilde{P}_{i}^{1}}[k]\vec{\theta}_{i}-\widehat{\tilde{\vec{s}}_{i}^{1}}[k]
\end{equation}

>>>>

% \loop\iftrue%
% \errmessage{This article is copyrighted by ___ and should not be TeXed}%
% \repeat%

@@

\section{INTRODUCTION}
\todo{Some citations:
  \cite{MaestreEtAl2021}

}
% % % \mpc{} \dmpc{}

@@

\begin{equation}
  \label{eq:constraint}
  \sum^{\nsubsystems}_{i=1}\Gamma_{i}\uik\preceq\umax
\end{equation}
where ${\Gamma_{i}:\R^{n_{u}\times n_{u}}}$\todo{$\Gamma_{i}:\semidefpos$?}, ${\umax:\R^{n_{u}\times 1}}$ and ${n_{u}=\card{\uik}}$.

\begin{problem}{Global \mpc{} Problem.}\label{Pb:GOP}
\begin{equation*}
\begin{matrix}
\underset{\useq}{\mathrm{minimize}}&\resizebox{0.35\textwidth}{!}{$\overbrace{\sum\limits^{M}_{i=1} \overbrace{\sum_{j=1}^{\predhorz} \norm{\mpcvec{v}[i][k+j][k]}^{2}_{Q_i}+\norm{\mpcvec{u}[i][k+j-1][k]}^{2}_{R_i}}^{\textstyle{} \obji}}^{\textstyle{} \globobj}$}\\
\mathrm{subject~ to}&~\eqref{eq:systems},\eqref{eq:upositive}\ \mathrm{and}\ \eqref{eq:constraint}
\left\}\small
\begin{aligned}
  &\forall i\in \{1:\nsubsystems\}\\
  &\forall j\in \{1:\predhorz\}
\end{aligned}\right.

@@

$\optglobobj$ denotes the optimal value of the problem~\ref{Pb:GOP}, and ${\optuseq}$ is the optimal control sequences for all agents.
The problem~\ref{Pb:GOP} is solved at each time $k$, and the ${\optuikk}$ are applied in each respective $i$ subsystem, following a receding horizon strategy.
\end{problem}

\subsection{Distributed Model Predictive Control}\label{ssec:dMPC}
\todo{\dmpc{},
}
\begin{subequations}
  \begin{equation}
    \left.
      \small
      \begin{aligned}
        J_{i}^{\star}(\thetaik)&=\underset{\uiseq}{\mathrm{minimize}} \obji\\
        \mathrm{s.t.} &\quad\eqref{eq:systems}\ \&\ \eqref{eq:upositive}\\
        &\quad\Gamma_{i}\uik\preceq\thetaik:\lambdaik\\
      \end{aligned}
    \right\}
    \small
    \begin{aligned}
      &\forall i\in \{1:\nsubsystems\}\\
      &\forall j\in \{1:\predhorz\}
    \end{aligned}
    \label{eq:DOP_local}
  \end{equation}
  \begin{equation}
    \small
    \begin{aligned}
      J^{\star}&=\underset{\thetaseq}{\mathrm{minimize}}\sum^{\nsubsystems}_{i=1} J^{\star}_i(\thetaik)\\
      \mathrm{s.t.} &\quad \sum_{i=1}^{\nsubsystems}\thetaik\preceq\umax
    \end{aligned}
    \label{eq:DOP_master}
  \end{equation}
\end{subequations}

@@

Where ${\set{H} = \setbuild{\vec{\theta}}{\sum_{i=1}^{M}\vec{\theta}_{i}=\vec{u}_{\max}}}$, $\vec{\theta}$ are the allocations for all agents.
\todo{The sum $\sum_{i=1}^{M}\vec{\theta}_{i}$ can also be represented by the matrix multiplication $I_{c}^{M}\vec{\theta}$. Where ${c=N_{p}\pi_{\uik}}$.}

\begin{equation}
  \label{eq:thetaNegot}
\thetai\pplusone=\thetai\p+\rho\left(\vec{\lambda}_{i}^{\star} (\vec{\theta}_{i}\p)-I_{c}^{M} \pseudoinv{I_{c}^{M}}\vec{\lambda}^{\star} (\vec{\theta}\p)\right)
\end{equation}

where ${I_{c}^{M}=\kron{\1_{M,1}}{I_{c}}}$.
\todo{Observe that ${I_{c}^{M} \pseudoinv{I_{c}^{M}}\vec{v}=\frac{\sum_{i=0}^{M}\vec{v}_{i}}{M}}$}

\todo{
  But as the coordinator must also assure that ${\thetaik\succeq\0}$ so the local problems are feasible, we set 0 as lower bounds:
  }
  \todo[verify if need to use max]{
\begin{equation}
  % \label{eq:thetaNegot}
  \thetai\pplusone=
  \left\{ \thetai\p+\rho\left(\vec{\lambda}_{i}^{\star} (\vec{\theta}_{i}\p)-I_{c}^{M} \pseudoinv{I_{c}^{M}}\vec{\lambda}^{\star} (\vec{\theta}\p)\right)
  ,0\right\}
\end{equation}
}
\no{\todo[delete probably]{
  But observe in~\eqref{eq:projectedSubgradient} that if ${\vec{\theta}_{i}\p=\0}$, ${(\vec{\lambda}_{i}^{\star} (\vec{\theta}_{i}\p)-I_{c}^{M} \pseudoinv{I_{c}^{M}}\vec{\lambda}^{\star} (\vec{\theta}\p))}$ needs to be negative so $\vec{\theta}_{i}\pplusone$ becomes negative, which is impossible \question{is it though?} since $\lambdaik\succeq\0$, from the KKT optimality conditions, seen in~\cite[Chapter 5]{BoydVandenberghe2004}.
}}

@@

\todo{As seen in~\cite{BemporadEtAl2002} and in~\cite{BorrelliEtAl2017}, the solution for a constrained linear quadratic regulation problem yields in a \pwa{} function.
Similarly, we can get an explicit  solution for its dual variables $\vec{\lambda}_{i}$, which is also a \pwa{} function with respect to $\vec{\theta}_{i}$:}
\begin{equation}
  \begin{aligned}
    \label{eq:lambdafuntheta}
    \lambdai=
    \begin{cases}
      -P_{i}^{1}\vec{\theta}_{i}-\vec{s}_{i}^{1}[k]&\text{if}\ G_{1}[k]\thetai \preceq z_{1}[k] \\
      \qquad\quad \vdots&\qquad\quad \vdots\\
      -P_{i}^{N}\vec{\theta}_{i}-\vec{s}_{i}^{N}[k]&\text{if}\ G_{N}[k]\thetai \preceq z_{N}[k] \\
    \end{cases}
  \end{aligned}
\end{equation}
The halfspaces defined by the set of pairs $(G_{j}[k],z_{j}[k])$ represent a combination of active constraints for a given time $k$.

\begin{challenge}\label{ch:partition_unknown}
  It is important to note that since the actual state is unknown we cannot anticipate the partition of the space.
\end{challenge}
\begin{challenge}\label{ch:zone_unknown}
 For the same reason the values of $\vec{s}_{i}[k]$ are also unknown.
\end{challenge}

 As each constraint can be active or inactive, for a group of $n$ constraints, we can have at most $N=2^{n}$ different combinations of active/inactive sets and consequently $N$ regions (or zones).
 \begin{remark}\label{re:N_different_zones}
  Observe that depending on the constraints, some of these combinations can redundant or even make the optimization infeasible.
 \end{remark}
 \begin{assumption}
   We assume that none of these cases in Remark~\ref{re:N_different_zones} can occur, and we have $N$ zones.
 \end{assumption}

@@

% \subsection{Detection and mitigation}
\begin{assumption}
  $\gamma_{i}(\cdot)$ is the same during the negotiation phase for a given time $k$ (it does not depend on $p$).
\end{assumption}

\begin{assumption}
  equation
  T inversivel
  We suppose the agent chooses a linear function such as \[ {\tilde{\lambdai}=\gamma_{i}(\vec{\lambda}_{i})=\Tik\vec{\lambda}_{i}} \] resulting in:
  \begin{equation}
    \label{eq:linear_cheating}
    \tilde{\lambdai}=
        \begin{cases}
      -\Tik P_{i}^{1}\vec{\theta}_{i}-\Tik \vec{s}_{i}^{1}[k]&\text{if}\ G_{1}[k]\thetai \preceq z_{1}[k] \\
      \qquad\quad \vdots&\qquad\quad \vdots\\
      -\Tik P_{i}^{N}\vec{\theta}_{i}-\Tik \vec{s}_{i}^{N}[k]&\text{if}\ G_{N}[k]\thetai \preceq z_{N}[k] \\
    \end{cases}
  \end{equation}
and we define $\tilde{P}_{i}^{j}[k]=\Tik P_{i}^{j}$ and $\tilde{\vec{s}}_{i}^{j}[k]=\Tik\vec{s}_{i}^{j}[k]$.
\end{assumption}
\todo{
  Since $\Theta_{i}$ are diagonal we can assure\pdfnote{proof?} there exists one \question{zone?}zone where all constraints are active. We will fix $j=1$ for this zone and called it the $1$-zones.
  We can estimate the values for this partition using:
}
\begin{equation}
  \label{eq:lambdafuntheta_tilde}
\tilde{\lambdai}=\gamma_{i}(\lambdai(\thetai))=-\widehat{\tilde{P}_{i}^{1}}[k]\vec{\theta}_{i}-\widehat{\tilde{\vec{s}}_{i}^{1}}[k]
\end{equation}
